---
title: "assignment4"
author: "Alexander Chase"
date: "10/15/2021"
output:
  pdf_document: default
  html_document: default
df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# More on Statistics

In this assignment you will further explore and discover one of the most important results in Statistics: The Central Limit Theorem for Proportions.

## Step 0

Start by loading some data into R. For this you will use the `runif` function to generate a sample of numbers from the uniform distribution (theoretical distribution where all observations from the population are equally likely).

1.  Look up how to use the `runif` function on the internet or by issuing the `?runif` in the R console.

2.  Use `runif` to produce 1000 random numbers in the interval from 0 to 20. Store this vector in a variable named `population`. To ensure reproducibility of your results set the seed of R's random number generator using the `set.seed` function.

3.  Visualize this sample's distribution and describe its shape. Use 20 bins.

```{r parts 1-3}
set.seed(123)
population = runif(1000, min= 0, max= 20)   # fix this line, must use runinf
hist(population,breaks = 20) # and this line by specifying the correct number of breaks
```

4.  Write a function `prop` that takes two arguments, a vector `samp` and a number `x`, that computes the proportion of values in `samp` that are greater than or equal to `x`.

5.  Use your `prop` function to find the proportion of numbers in `population` that are greater than or equal to 12. Store this in a variable named `true_proportion`. Report its value.

```{r code for parts 4-5}

prop = function(samp,x){
    clump <- vector()
  {for (c in samp)
     if (c >= x) { 
      clump <- c(clump, c)
    }
    
 return (length(clump)/length(samp))
print(prop)
}}
# Run these tests to make your function works
stopifnot(prop(seq(1,10),5.5)==0.5)
stopifnot(prop(seq(1,1000),500.5) == 0.5)
true_proportion = prop(population, 12) # fix this line to compute the true proportion in population
print(true_proportion)
```

## Step 1

Let us pretend that we don't know the value of `true_proportion` and that we don't have access to all the values in `population`.

6.  Think about this for a minute: How would you estimate the value of `true_proportion`? Now go to part 7.

7.  To estimate `true_proportion` we could take samples from `population`. Let's do that! We could take samples of size 15. Randomly select 15 values from `population`. Use your `prop` function to obtain the proportion of these numbers that are greater than or equal to 12.

```{r first sample of size 15}
# Your code goes here for 6-7 goes here
set.seed(321)
sample1 = sample(population,size=15)
estimate1<-prop(sample1,12)
print(estimate1)

```

8.  Compare your first estimate `estimate1` to the true proportion `true_proportion` using the *relative error*: $\frac{\text{approximation}}{\text{true value}}-1.$ Drop the sign in case you get a negative number (using absolute value).

```{r relative error}

print(abs(estimate1/true_proportion-1)) # fix this line to compute the relative error for your estimate

```

9.  A relative error greater than 0.05 (5%) is large. This will depend on how large your quantities are. Maybe we will have better luck with a different sample. Try again! Randomly select another 15 values from `population`. As before, use `prop` to obtain the proportion of these numbers that are greater than or equal to 12.

```{r second sample of size 15}
set.seed(456)
sample2 = sample(population, size=15)
estimate2 <- prop(sample1, 12)
print(estimate2)
```

10. Compute the relative error for your second estimate. Did you have better luck this time?

```{r relative error for second sample}
print(abs((estimate2/true_proportion)-1))
    
```

11. Let us try this many times (Law of Large Numbers!). Let us take 1000 samples, their respective proportions, and average them out (this is your new proportion `estimate`). Report your result.

```{r}
set.seed(324)
samples = replicate(1000,prop(sample(population,15),12)) # fix this line by filling in the blanks
estimate = mean(samples)
print(estimate)
```

12. Are you starting to get lucky? Compute the relative error for your `estimate`. How good is it in terms of its relative error?

13. Visualize `samples`. Describe its shape.

```{r sampling dist results}
print(abs((estimate/true_proportion)-1))
hist(samples)
```

14. Is your estimate better? Do it again (parts 11-13) for good measure (be sure to change the seed value).

```{r}
set.seed(983)
sample3 = replicate(1000, prop(sample(population, size= 15), 12))
estimate3 = mean(sample3)
hist(sample3)
print(abs((estimate3/true_proportion)-1))
```

## Step 3

15. What happens if you take larger samples? Repeat parts 10-13 using samples of size 25 with 10000 replications. Report what happened.

```{r}
set.seed(333)
sample4 = replicate(10000, prop(sample(population, size= 25), 12))
estimate4 = mean(sample4)
hist(sample4)
print(abs((estimate4/true_proportion)-1))
```
